---
title: "Health Conditions"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    theme: bootstrap
    navbar:
       - { title: "Explore Economic Investment", href: property_values.html, align: left }
       - { title: "Statistical Analysis", href: analysis.html, align: left }
       - { title: "Back to Home", href: index.html, align: right }

---

```{r}

library(tidyverse)
library(plotly)
library(leaflet)
library(ggridges)
library(rgdal)
library(sp)

#Set global theme for plots
theme_set(theme_minimal() + 
            theme(legend.position = "bottom", 
                  axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)))
#Set options for all plots
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d


```

```{r, message = FALSE, warning = FALSE, include = FALSE}
health_data_final <-
  read_csv("./data/final_df.csv") %>% 
  select(community_board, borough, obesity, hypertension, life_expectancy, self_rep_health, garden_num, id_spatial) #%>% 
  
working <- getwd()
com_board_spdf <- readOGR(dsn = working, layer = "community_board_new")
names(com_board_spdf)
health_data_spatial <- merge(com_board_spdf, health_data_final, by.x = "boro_cd", by.y = "id_spatial")

```


Column {data-width=400}
-----------------------------------------------------------------------

### Community Gardens

<style type="text/css">

.chart-title {  /* Community Gardens  */
   font-size: 20px;

</style>


```{r}
# Create a color palette for the map:
pal <- colorNumeric("YlGn", health_data_spatial@data$garden_num)

# Prepare the text for labels:
mytext <- paste(
    "Borough: ", health_data_spatial@data$borough,"<br/>",
    "Community Board: ", health_data_spatial@data$community_board,"<br/>", 
    "Number of Gardens: ", health_data_spatial@data$garden_num, "<br/>", 
    sep = "") %>%
  lapply(htmltools::HTML)
 
# Final Map with shaded colors by number of gardens in the community district
leaflet(health_data_spatial) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%  
  setView( lat = 40.7, lng = -74 , zoom = 10) %>%
  addPolygons( 
    fillOpacity = 0.8, 
    smoothFactor = 0.5, 
    fillColor = ~colorQuantile("YlGn", garden_num)(garden_num), 
    stroke = TRUE, 
    color = "black", 
    weight = 0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal = pal, values = ~garden_num, opacity = 0.9, title = "Number of Gardens", position = "topleft" )
```

> Visualization of the distribution of number of gardens across community boards.

Column {.tabset .tabset-fade data-width=400}
-----------------------------------------------------------------------

### Obesity


```{r}
# Create a color palette for the map:
pal2 <- colorNumeric("BuPu", health_data_spatial@data$obesity)

# Prepare the text for labels:
mytext2 <- paste(
  "Borough: ", health_data_spatial@data$borough,"<br/>",
    "Community Board: ", health_data_spatial@data$community_board,"<br/>", 
    "Percent Obesity: ", health_data_spatial@data$obesity, "<br/>", 
    "Number of Gardens: ", health_data_spatial@data$garden_num, "<br/>",
    sep = "") %>%
  lapply(htmltools::HTML)
 
# Final Map with shaded colors by percent obesity in the community district
leaflet(health_data_spatial) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%  
  setView( lat = 40.7, lng = -74 , zoom = 10) %>%
  addPolygons( 
    fillOpacity = 0.8, 
    smoothFactor = 0.5, 
    fillColor = ~colorQuantile("BuPu", obesity)(obesity), 
    stroke = TRUE, 
    color = "black", 
    weight = 0.3,
    label = mytext2,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal = pal2, values = ~obesity, opacity = 0.9, title = "Percent Obesity", position = "topleft" )
```

> Visualization of the distribution of obesity across community boards. 
> Obesity: Percentage of adults ages 18 and older who have obesity (Body Mass Index of 30 or greater) based on self-reported height and weight.

### Hypertension


```{r}
# Create a color palette for the map:
pal3 <- colorNumeric("BuPu", health_data_spatial@data$hypertension)

# Prepare the text for labels:
mytext3 <- paste(
    "Borough: ", health_data_spatial@data$borough,"<br/>",
    "Community Board: ", health_data_spatial@data$community_board,"<br/>", 
    "Percent Hypertension: ", health_data_spatial@data$hypertension, "<br/>", 
    "Number of Gardens: ", health_data_spatial@data$garden_num, "<br/>",
    sep = "") %>%
  lapply(htmltools::HTML)
 
# Final Map with shaded colors by percent hypertension in the community district
leaflet(health_data_spatial) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%  
  setView( lat = 40.7, lng = -74 , zoom = 10) %>%
  addPolygons( 
    fillOpacity = 0.8, 
    smoothFactor = 0.5, 
    fillColor = ~colorQuantile("BuPu", hypertension)(hypertension), 
    stroke = TRUE, 
    color = "black", 
    weight = 0.3,
    label = mytext3,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal = pal3, values = ~hypertension, opacity = 0.9, title = "Percent Hypertension", position = "topleft" )
```

> Visualization of the distribution of hypertensio across community boards. 
> Hypertension: Percentage of adults ages 18 and older who report ever being told by a healthcare professional that they have hypertension, also known as high blood pressure.


### Life Expectancy



```{r}
# Create a color palette for the map:
pal4 <- colorNumeric("BuPu", health_data_spatial@data$life_expectancy)

# Prepare the text for labels:
mytext4 <- paste(
    "Borough: ", health_data_spatial@data$borough,"<br/>",
    "Community Board: ", health_data_spatial@data$community_board,"<br/>", 
    "Life Expectancy: ", health_data_spatial@data$life_expectancy, "<br/>", 
    "Number of Gardens: ", health_data_spatial@data$garden_num, "<br/>",
    sep = "") %>%
  lapply(htmltools::HTML)
 
# Final Map with shaded colors by life expectancy in the community district
leaflet(health_data_spatial) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%  
  setView( lat = 40.7, lng = -74 , zoom = 10) %>%
  addPolygons( 
    fillOpacity = 0.8, 
    smoothFactor = 0.5, 
    fillColor = ~colorQuantile("BuPu", life_expectancy)(life_expectancy), 
    stroke = TRUE, 
    color = "black", 
    weight = 0.3,
    label = mytext4,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal = pal4, values = ~life_expectancy, opacity = 0.9, title = "Life Expectancy", position = "topleft" )
```

> Visualization of the distribution of life expectancy across community boards. 
> Life expectancy: Life expectancy at birth.


### Self Reported Health

```{r}
# Create a color palette for the map:
pal5 <- colorNumeric("BuPu", health_data_spatial@data$self_rep_health)

# Prepare the text for labels:
mytext5 <- paste(
    "Borough: ", health_data_spatial@data$borough,"<br/>",
    "Community Board: ", health_data_spatial@data$community_board,"<br/>", 
    "Self Reported Health: ", health_data_spatial@data$self_rep_health, "<br/>",
    "Number of Gardens: ", health_data_spatial@data$garden_num, "<br/>",
    sep = "") %>%
  lapply(htmltools::HTML)
 
# Final Map with shaded colors by self reported health in the community district
leaflet(health_data_spatial) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%  
  setView( lat = 40.7, lng = -74 , zoom = 10) %>%
  addPolygons( 
    fillOpacity = 0.8, 
    smoothFactor = 0.5, 
    fillColor = ~colorQuantile("BuPu", self_rep_health)(self_rep_health), 
    stroke = TRUE, 
    color = "black", 
    weight = 0.3,
    label = mytext5,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal = pal5, values = ~self_rep_health, opacity = 0.9, title = "Self Reported Health", position = "topleft" )
```

> Visualization of the distribution of self reported health across community boards. 
> Self Reported Health: Percentage of adults ages 18 and older who report their overall health is “excellent,” "very good" or “good” on a scale of excellent, very good, good, fair or poor.




